---
title: "MUDAC 2019: Minnesota Water Quality"
author: "Genevieve Block, Ben Christensen, Sara Lawson, Delia Matos"
date: "April 23, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(readxl)
require(knitr)
require(png)
require(GGally)
```


```{r, echo = FALSE}
LSueur <- read_csv("les.csv")[1:6]
Root <- read_csv("root.csv")
Otter <- read_csv("otter.csv")
StLou <- read_csv("stlou.csv")[1:6]
LinMod <- read_csv("Linear Model Data.csv")
names(LSueur)[2]<-"Rain/Snow"
names(Root)[3]<-"Timestamp"
names(LSueur)[2] <- "Rain"
names(Root)[2] <- "Rain"
names(Otter)[2] <- "Rain"
names(StLou)[2] <- "Rain"
```

```{r, echo = FALSE}
LSueur %>%  
  mutate(watershed = "LS") -> LSueur
Root %>%  
  mutate(watershed = "R") -> Root
Otter %>%  
  mutate(watershed = "O") -> Otter
StLou %>%  
  mutate(watershed = "St") -> StLou
```


```{r, echo = FALSE}
alldata <- rbind(LSueur, Root, Otter, StLou)
```

```{r, echo = FALSE}
allDataNoZeroes <- alldata %>% filter(Intensity > 0) %>% filter(Rain != "Snow")
rootDataNoZeroes <- allDataNoZeroes %>% filter(watershed == "R") %>% filter(Rain != "Snow")
otterDataNoZeroes <- allDataNoZeroes %>% filter(watershed == "O") %>% filter(Rain != "Snow")
stLouDataNoZeroes <- allDataNoZeroes %>% filter(watershed == "St") %>% filter(Rain != "Snow")
lSueurDataNoZeroes <- allDataNoZeroes %>% filter(watershed == "LS") %>% filter(Rain != "Snow")
```


## Outline


## Problem 1

- The question

Problem: find the characteristics of watersheds that effect the river water quality of Minnesota.  

```{r fig.width=1.5, fig.height=1.5,echo=FALSE}
library(png)
library(grid)
img <- readPNG("All_watersheds.PNG")
 grid.raster(img)
```

- Minnesota Pollution Control Agency; Water Resource Center at Minnesota State University Mankato

- Data size
50 observations with 19 variables 

- Goal: prediction v inference; regression v classification
regression vs. classification

## The Data

Dependent variables:
 - Nitrate
 - Total Suspended Solid (TSS)

Independent variables:
 - Forest
 - Grass and Hay
 - Wetlands
 - Cropland
 - Developed/Urban
 - Altered Streams
 - Tile-drained land
 - Shallow Bedrock under cropland
 - Sand
 - Silt
 - Clay
 - Organic Matter
 - Land Slope
 - Lakes
 - Lake interception
 - Latitude
 - Longitude 

## Results

From our model we found that Nitrate and TSS levels are influenced by:
 - Altered Streams (Nitrate & TSS)
 - Tile-drained land (Nitrate & TSS)
 - Shallow Bedrock under cropland (Nitrate & TSS)
 - Silt (Nitrate & TSS)
 - Latitude (Nitrate & TSS)
 - Cropland (TSS)
 - Sand (TSS)
 - Clay (TSS)
 - Lake interception (TSS)
 - Nitrate (TSS)
 
Conclusions 

From our analysis, we believe that the state of Minnesota should:
 - Create more policy for agricultural tile drainage
 - Filter agricultural runoff with constructed and restored wetlands


## Data Cleaning

- Renaming variables
- Didn't have to change variable types 



## Data Exploration
- ggcorr plotto see the correlations between all the variables 
- ggpairs to show the different distributions of the variables 

```{r Data Exploration}
ggcorr(DataNew, palette = "RdBu", label = TRUE)
ggpairs(DataNew)
```


## Cross-Validation


## Variable Selection

```{r}
BestNitrate <- regsubsets(Nitrate ~ ., DataNew, nvmax = 18)
summary(BestNitrate)
BestSumNitrate <- summary(BestNitrate)
```

```{r}
BestTSS <- regsubsets(`Total Suspended Solid (TSS)` ~ ., DataNew, nvmax = 18)
summary(BestTSS)
BestSumTSS <- summary(BestTSS)
```
```{r}
plot(BestSumNitrate$adjr2)
plot(BestSumTSS$adjr2)
```

```{r}
coef(BestNitrate, 10)
coef(BestTSS, 10)
```
## Model Analysis

- What do the statistics and visualizations tell you about the strength, efficacy, accuracy of your model?


## Summary

Our process:
- Load data
- Explore data
- Subset 
- Cross Validation 
- Regressions
- Reidual sqaured error 

Conclusion: 
- The State of Minnesota should monitor their farming waste policies and create wetlands to counteract the agricultural run-off.
 
 
## Problem 2

- What is the effect of rainfall intensity on water quality?

        - Create Annual Precipitation Intensity Index
        - Measure correlation between index and TSS, Nitrate

- Minnesota Pollution Control Agency; Water Resource Center at Minnesota State University Mankato

- 2,875,000 rows of hourly precipitation data

        - 56 precipitation zones
        - 4 watersheds
        - 8 years

- Unsupervised; inference; regression (quantitative variables)

## Watershed Map


```{r fig.width=4.5, fig.height=4.5,echo=FALSE}
library(png)
library(grid)
img <- readPNG("4_watersheds.PNG")
 grid.raster(img)
```


## The Data

- Location (data type)

- Timestamp (data type)

- Intensity (data type)

- Various year ranges 

        - 2007 - 2012
        - 2010 - 2014
        - 2009 - 2014


## Results

- Rainfall intensity does influence TSS

- Rainfall intensity does not influence Nitrates

- Our index is a reasonable metric of TSS


## Data Cleaning

- Convert to flat format in Excel

- Add columns 
        
        - rain/snow/spring (character)
        - Datetime (numeric)

- rbind to create one dataset


## Data Exploration

- intensity ranges

- outliers

- bins to view distribution of intensities 

- INSERT GRAPHIC OF BINS OR R CODE


## Variable Selection

- Heuristic from Problem Statement

- What information can we get out of the data? What variables can we create for the index?

- Which index variables contribute more?

- APII = .5 * (Hourly Rainfall Points) + .5 * (Maximum Points)


## Model Analysis

- We'll use lm(Index ~ TSS... to do this

        - for each watershed and over all years
        - Emphasize R^2

- Show graphs


```{r, echo = FALSE}
APII_TSS_LS <- lm(Index ~ TSS, data=subset(LinMod, Watershed == "Le Sueur") )

summary(APII_TSS_LS)

APII_TSS_Otter <- lm(Index ~ TSS, data=subset(LinMod, Watershed == "Otter Tail") )

summary(APII_TSS_Otter)

APII_TSS_Root <- lm(Index ~ TSS, data=subset(LinMod, Watershed == "Root") )

summary(APII_TSS_Root)

APII_TSS_SL <- lm(Index ~ TSS, data=subset(LinMod, Watershed == "St. Louis") )

summary(APII_TSS_SL)

APII_Nitrate_LS <- lm(Index ~ Nitrate, data=subset(LinMod, Watershed == "Le Sueur"))

summary(APII_Nitrate_LS)

APII_Nitrate_Otter <- lm(Index ~ Nitrate, data=subset(LinMod, Watershed == "Otter Tail") )

summary(APII_Nitrate_Otter)

APII_Nitrate_Root <- lm(Index ~ Nitrate, data=subset(LinMod, Watershed == "Root") )

summary(APII_Nitrate_Root)

APII_Nitrate_SL <- lm(Index ~ Nitrate, data=subset(LinMod, Watershed == "St. Louis") )

summary(APII_Nitrate_SL)
```


## Summary

 - Rainfall intensity can be measured with some success using total amount of rainfall over period (weighted by time of year) and outliers
 
 - A strong APII could be used as another predictor in Problem 1
 


# MUDAC 2019: Minnesota Water Quality 